Contextual Identity: A Framework for Usable Privacy

### 2012-10-13 paper.tex
Squashed commit of the following:
commit 18beeeec53ffb809836818d695e0c8d5ad63e59a
Author: Sid Stamm <sid@mozilla.com>
Date:   Fri Oct 12 17:37:30 2012 -0700

    Added direct use of the formal model.  Also added stuff about google sharing and tor.

We attempt to formalize the notion of contextual identity, use it as a
framework to understand existing privacy violations, and propose mitigations to
those violations.

@@

\section{Previous work}

Barth et al.~provide a formal model for contextual integrity of user
data~\cite{barth}. The contextual integrity model focusses on the flow of
personal information in terms of knowledge states held by agents rather than on
user identity, and works well for modeling privacy legislation and policy. In
contrast, this work's model provides a framework to understand users' ability
to represent which identity they want to use depending on context, and works
well for modeling software features.

\begin{comment}TODO: figure out how to cite this. Why is this short
paper's model worth pursuing? It's simpler, and focusses on the user rather
than data.
\end{comment}

Although network attacks are very powerful, at the current time they require a
level of expertise and resources not typically held by those causing the most
worry to users. A brief survey of social networking advice reveals that the
major causes of concern come not from a dedicated graph expert, but people that
the user knows: friends, family, and co-workers~\cite{fbtips2,fbtips1}.
Surveys conducted by Wang et al.~show that typical regrets from posting on
Facebook stem almost exclusively from fear of negative interactions with people
that the users knows~\cite{wang}. The consequences of these negative
interactions can lead to loss of employment or breaking personal relationships.
By contrast, much privacy research has focussed on network privacy, where the
bad actor is a behavioral tracking service, or a state agent.

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts\footnote{Google Talk contacts are everyone
with whom a user has chatted, which might include co-workers, supervisors, and
friends}. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts (which users used to treat as private
messages, before the advent of ``Like'' and comment buttons) to the attention
of an audience that the user never intended~\cite{timeline}.

@@

\subsection{Unsatisfiable policies}
Sometimes service providers have policies which preclude isolating multiple
identites. For example, Facebook and Google have a ``Real Names'' policy, which
requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies presume the users can be one
person to everyone. These policies ignore that community-building happens in
many different contexts, that users have many legitimate reasons for presenting
different identities in different contexts, and that users don't necessarily
want those identities to be linked. For example, disallowing avatar handles as
a primary identifier makes building a gaming community difficult at best, and
linking a gaming identity to a professional identity has already caused
problems for one exposed has already caused problems for Maine Senate candidate
and World of Warcraft gamer Colleen Lachowicz in 2012~\cite{maine}.

It is impossible for users who want to isolate multiple contextual identities
to participate in these networks without violating the terms of service.

\subsection{Federated login}
Facebook Connect is a login platform that allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}. Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities.

Other login platforms and protocols, including OAuth and BrowserID
do not suffer from this policy or design error~\cite{browserid,oauth}.
\begin{comment}
(TODO:
this is actually kinda complicated -- yahoo, twitter, google allow multiple
accounts but the multi-login support is broken at google, not sure how it is
with others, oauth protocol supports multiple accounts, persona clearly
supports multiple accounts)
\end{comment}

\subsection{User interface errors}
Even in the absence of changes made by service providers to re-contextualize
personal information, it is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently sent compromising pictures of
himself publicly~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{Third party tracking}
\label{sec:tracking}
Even if a user does not participate in social networks, they are still subject
to tracking by virtue of browsing the web. Third-party cookies, flash cookies,
web bugs, device fingerprinting, geolocation, history sniffing and more are
techniques that a dedicated third party could use to collect browsing history
and link contextual identities.

@@

\subsection{Automatic mitigations for information leakage}
Given the set of relationships $R$ and tasks $T$, we can categorize a user's
contextual identities and prevent information leakage between them.  For
example, preventing third-party cookies from being read and set across task
boundaries means that targeted ads for dating would not follow the user when
visiting a technology news site. Knowing the user's current contextual identity
also allows the browser to suppress URL bar suggestions that aren't relevant to
that identity, protecting the user from both shoulder-surfing and distraction.

Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, $R$ when
using private browsing mode may be other household members, or it may be a
service to which the user wants to limit cookie leakage. Another deficit is
that private browsing mode does not adequately serve long-term tasks (such as
curating a set of adult entertainment sites) in any major browser.

Another good partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}. In this case, when the
user decides to register for a service using Persona, the task $T$ they are
performing is authentication, the relationships $R$ are ones they will engage
in when using the service, and $p$ is tied to the email identifier they use to
register for the service. Because we can categorize the type of service by URL
as \texttt{aboutprofile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

@@

\subsection{Recovering from mistakes}
\subsubsection{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in the
perfect position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, topics that are
frequently regretted~\cite{wang}. Although this approach does not help when the
service provider decides to change the audience on the user's behalf, auditing
gives the user a good opportunity to review posts that might cause
regrets. In the case of high-volume users, such a tool could even incorporate
sentiment analysis to aid the user in finding particularly problematic content.

Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via techniques in~\ref{sec:tracking}. Collusion is a
tool for visualizing cookies~\cite{collusion}, in particular the third-party
cookies set by tracking sites which are typically ad networks.  Once the user
knows which sites are tracking them, they can choose to block cookies from
those sites.

\subsubsection{Remembering to forget}
Humans don't have perfect memories, and neither should social
networks~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily. Both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

@@

\subsection{Preventing mistakes}
\subsubsection{Slowing information flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.

Protocols like Do Not Track are also a good first step to solving this
problem~\cite{dnt}. Although Do Not Track is unenforceable from the client
side, it provides a clear signal to service providers and ad networks
that the user does not want to be tracked. As adoption increases, ignoring this
signal will lead to loss of reputation for service providers that explicitly
act against their users' wishes.

@@

\begin{comment}
\section{Acknowledgements}
The authors thank
Ben Adida
Ed Lee
Lucas Adamski
Mike Connor

\end{comment}

>>>>

We provide a simple formalization of the notion of contextual identity, use it as a
framework to understand existing privacy violations, and analyze mitigations to
those violations.

\section{Related Work}

Barth et al.~provide a formal model for contextual integrity of user
data~\cite{barth}. The contextual integrity model focuses on the flow of
personal information in terms of knowledge states held by agents rather than on
user identity, and works well for modeling privacy legislation and policy. In
contrast, this work's model provides a simple framework to understand users' ability
to represent which identity they want to project depending on context, and works
well for modeling how software features affect users' ability to assert choice
and control.

\begin{comment}TODO: figure out how to cite this. Why is this short
paper's model worth pursuing? It's simpler, and focuses on the user rather
than data.
\end{comment}

Although network attacks are very powerful, at the current time they require a
level of expertise and resources not typically held by those causing the most
worry to users. A brief survey of social networking advice reveals that the
major causes of concern come not from a dedicated graph expert, but people
close to the user: friends, family, and co-workers~\cite{fbtips2,fbtips1}.
Surveys conducted by Wang et al.~show that typical regrets from posting on
Facebook stem almost exclusively from fear of negative interactions with people
that the users knows~\cite{wang}. The consequences of these negative
interactions can lead to loss of employment or breaking personal relationships.
Most previous privacy research has focused on distant adversaries, where the
bad actor is a behavioral tracking service, state agent or unknown
eavesdropper; in contrast, we assert that users care more about their
interactions with others close to them.

@@

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
(Lack of control over $R$).\\
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts\footnote{Google Talk contacts are everyone
with whom a user has chatted, which might include co-workers, supervisors, and
friends}. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts (which users used to treat as private
messages, before the advent of ``Like'' and comment buttons) to the attention
of an audience that the user never intended~\cite{timeline}.

@@

\subsection{Unsatisfiable policies}
(Lack of control over $p$).\\
Sometimes service providers have policies which preclude isolating multiple
identites. For example, Facebook and Google have a ``Real Names'' policy, which
requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies presume the users can be one
person to everyone. These policies ignore that community-building happens in
many different contexts, that users have many legitimate reasons for presenting
different identities in different contexts, and that users don't necessarily
want those identities to be linked. For example, disallowing avatar handles as
a primary identifier makes building a gaming community difficult at best, and
linking a gaming identity to a professional identity has already caused
problems for one exposed has already caused problems for Maine Senate candidate
and World of Warcraft gamer Colleen Lachowicz in 2012~\cite{maine}.

It is impossible for users who want to isolate multiple contextual identities
to participate in these networks without violating the terms of service.  The
Google ``legal name'' policy prohibits users from asserting control over $p$.

\subsection{Federated login}
(Lack of flexibility in $p$).\\
Facebook Connect is a login platform that allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}. Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities.

Similar to the Unsatisfiable policies, many federated login systems simply
don't have the ability to allow users multiple $p$ values, or identifiers.
Other federated login systems may support and allow multiple identifiers, but
it can be poorly implemented or difficult to discover, which essentially means
the feature isn't used.
There are some login platforms and protocols, including OAuth and BrowserID
that do not suffer from this policy or design error~\cite{browserid,oauth},
giving a user better control over which value of $p$ he employs in his
contextual identity.
\begin{comment} (TODO:
this is actually kinda complicated -- yahoo, twitter, google allow multiple
accounts but the multi-login support is broken at google, not sure how it is
with others, oauth protocol supports multiple accounts, persona clearly
supports multiple accounts)
\end{comment}

\subsection{User interface errors}
(Lack of control over $R$).\\
Even in the absence of changes made by service providers to re-contextualize
personal information, it is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently sent compromising pictures of
himself publicly~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{Third party tracking}
\label{sec:tracking}
(Lack of control over $T$, and correlation across values of $p$).\\
Even if a user does not participate in social networks, they are still subject
to tracking by virtue of browsing the web. Third-party cookies, flash cookies,
web bugs, device fingerprinting, geolocation, history sniffing and more are
techniques that a dedicated third party could use to collect browsing history
and link contextual identities through correlation of tasks in $T$ for various
identity values $p$.

@@

\subsection{Automatic mitigations for information leakage}
Given a set of relationships $R$ and tasks $T$, one can categorize a user's
contextual identities and prevent information leakage between them.  For
example, preventing third-party cookies from being read and set across task
boundaries means that targeted ads for dating would not follow the user when
visiting a technology news site. Knowing the user's current contextual identity
also allows the browser to suppress URL bar suggestions that aren't relevant to
that identity, protecting the user from both shoulder-surfing and distraction.

Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, $R$ when
using private browsing mode may be other household members, or it may be a
service to which the user wants to limit cookie leakage. A strong contextual
identity mechanism allows users control over $R$ or over $T$ when $R$ is
clearly identified to them, but the various privacy modes in different web
browsers either allows the user control over $T$ or over $R$, but not both at
the same time.

A partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}. In this case, when the
user decides to register for a service using Persona, the task $T$ they are
performing is authentication, the relationships $R$ are ones they will engage
in when using the service, and $p$ is tied to the email identifier they use to
register for the service. Because we can categorize the type of service by URL
as \texttt{aboutprofile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.  Persona allows the user direct
control over $p$ given a context ($R,T$) in which they're authenticating.

@@

\subsection{Recovering from Mistakes}
\subsubsection{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in Section~\ref{sec:examples}. A browser is in an
ideal position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.

For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, topics that are
frequently regretted~\cite{wang}.  This redacting is equivalent to asserting
better control over $T$. Although this approach does not help when the service
provider decides to change the audience on the user's behalf (loss of control
over $R$), auditing gives the user a good opportunity to review posts that
might cause regrets. In the case of high-volume users, such a tool could
incorporate sentiment analysis to aid the user in finding particularly
problematic content.  This effectively helps users adjust $T$ when $R$ changes
unexpectedly; while it does not help bring users into better control over $R$,
it can help them minimize the impact of a changing $R$.

Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via techniques in Section~\ref{sec:tracking}. Collusion is a
tool for visualizing cookies~\cite{collusion}, in particular the third-party
cookies set by tracking sites which are typically ad networks.  Once the user
knows which sites are tracking them, they can choose to block cookies from
those sites.  This type of tool allows users good control over $R$ by limiting
who sees the tracks they create.

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and one could argue that social
networks should also forget~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.
This would help users better contain the whole context ($R,T$) by segmenting
one large audience and one large set of tasks into smaller temporal partitions.

@@

\subsection{Preventing Mistakes}
\subsubsection{Controlling Information Flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.
This type of tool (cookie blocker) helps users assert control over audience ($R$).

Protocols like Do Not Track are also a good first step to solving this
problem~\cite{dnt}. Although Do Not Track is unenforceable from the client
side, it provides a clear signal to service providers and ad networks
that the user does not want to be tracked. As adoption increases, ignoring this
signal will lead to loss of reputation for service providers that explicitly
act against their users' wishes.
This type of tool (cookie blocker) helps users assert control over the effects of their actions ($T$).

The Google Sharing project allows users of Google services to share and mix the way Google identifies them~\cite{googlesharing}.  The effect is that users employ a contextual identity $\{T_{gs}, R_{gs}, p\}$ with a shared $p$.  These users employ it only for Google sites ($R_{gs} = Google$) and can accomplish tasks $T_{gs}$ while associating all their actions with a random $p$ also associated with many other users.  This devalues $p$ and makes it difficult to correlate $T_{gs}$ with $T$ from another of the users' contextual identities.

Channel encryption systems like TOR\footnote{The Onion Router: \url{https://www.torproject.org}} allow the users more network secrecy; this allows them to better control their audience $R$ by hiding their actions from network operators.

@@

\begin{comment}
\section{Acknowledgements}
The authors thank
Lucas Adamski
Ben Adida
Mike Connor
Ed Lee
\end{comment}
